---
title: "SWITCHING IT UP: MIXED FEEDING PLAN MAY LEAD TO IMPROVED UTILIZATION OF DIETARY PLANT PROTEIN"
author: "Aubrey Dissinger"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: 
---

[Github](https://github.com/aubreydissinger/final-project.git)

## Introduction

Increasingly popular as a fishmeal (FM) replacement in aquafeeds, soybean meal (SBM) is widely available, cost effective, and high in protein (Park et al. 2017, Lin and Luo 2011, Zhou et al. 2005). However, there are negative attributes to dietary SBM due to the presence of several anti-nutritional factors (Makkar 1993, Francis et al. 2001). These negative effects can present as reduced feed intake, reduced growth performance, and the development of SBM induced enteritis (SBMIE; Hua et al. 2019, Francis et al. 2001, Zhou et al. 2005, Krogdahl et al. 2010). Typical histological signs of SBMIE include shortening of mucosal folds, thickening of lamina propria, and infiltration of inflammatory cells (Gu et al. 2016, Fuentes-Quesada et al. 2018, Agboola et al. 2022a), all of which interfere with the ability of the intestine to efficiently digest and absorb nutrients.

Past studies have shown varying time frames for the onset of enteritis symptoms. One study of grass carp (*Ctenopharyngodon idellus*), a herbivorous species, found the symptomatic stage of SBMIE to occur between 7 to 10 days of feeding with a 70% SBM diet (Peng et al. 2020). In omnivorous species, the onset may occur earlier for lower levels of SBM inclusion. A study of the common carp (*Cyprinus carpio*) found the first visible morphological evidence of enteritis after 7 days of feeding with a 20% SBM diet (Urán et al. 2008). In the case of zebrafish (*Danio rerio*), signs of intestinal inflammation have arisen between days 4 (Hedrera et al. 2013) and 10 (Rehmann et al. 2022) when fed 50% SBM diets. Carnivorous species are less equipped to consume plant feedstuffs (Oliva-Teles et al. 2015) and signs of intestinal inflammation present at lower inclusion levels of SBM (Sales 2009). Morphological changes in the intestine were detected in Atlantic salmon (*Salmo salar*) after 5 days of feeding with a 20% SBM diet and 7 days of feeding with a 10% SBM diet (Urán et al. 2009).

While the onset of SBMIE varies between species and feeding habit, it typically takes a few days of consistent SBM diet feeding to occur. Thus, varying the feeding regimen and utilizing a pattern of alternating diets may allow for the intestinal tract to heal or cope with the anti-nutritional factors present in SBM. Several studies have investigated the effects of feeding regimen on growth (Lawrence et al. 2012, Gonzales and Law 2013, Farias and Certal 2016), reproduction (Lawrence et al. 2012, Gonzales and Law 2013, Farias and Certal 2016), metabolism (Fang et al. 2013), and behavior (Dametto et al. 2018). However, not much is known on how varying feeding regimens affect SBM utilization and SBMIE. Thus, the objective of this study was to use the zebrafish, an omnivorous species with a middle-ground tolerance to SBM (Hedrera et al. 2013), as a model to investigate the effects of alternating feeding regimen with SBM diet on growth performance, feed utilization, and intestinal health.

## Methods

**Experimental Design and Feeding Regimen**

Twelve 3.0 L tanks were assigned in triplicate (n = 3) to one of four treatment groups: fish that received the same FM-based diet every day (**FM-FM**), fish that received the same SBM-based diet every day (**SBM-SBM**), fish that received alternating FM- and SBM-based diets daily (**FM-SBM**), and fish that received alternating FM- and SBM-based diets weekly (**FM/SBM**). At 22 days-post-fertilization (dpf), zebrafish were randomly stocked into treatment tanks one fish at a time until all tanks contained 21 fish. Thirty initial fish samples were euthanized using an ice slurry of equal parts ice and water (Wallace et al. 2018), patted dry, and weighed to obtain an average initial weight. This average initial weight was used to calculate the starting biomass of each tank. The feeding trial was carried out from 23 dpf to 49 dpf, before zebrafish reached sexual maturation. This was done to prevent any biases in growth performance due to sexual dimorphism.

To avoid differences in experimental diet feed intake resulting from stress-related anorexia, the feeding trial began 24 hours after stocking at 23 dpf. Zebrafish were fed their respective diets to apparent satiation three times per day. The feeding periods were 09:30-10:30, 13:00-14:00, and 16:30-17:30. Due to the varying feeding regimens between tanks, care was taken to create a satiation feeding protocol that limited biases between treatments as well as negligible feed left uneaten. Within each feeding period, each tank was fed in three separate rounds. During each round, zebrafish were offered a small amount of food, and their feeding behavior was carefully observed. Tanks were provided additional small amounts of feed if “hunting” behavior was observed, and fish were monitored to ensure these offerings were consumed. This was repeated five minutes later and again ten minutes later. Overall, each tank had three carefully observed fifteen-minute feeding periods per day. For each tank, feed offerings were halted once signs of slowed feeding behavior in each tank were observed. More specifically, feeding ceased when only two fish remained feeding in the tank or when the feeding period ended.

**Sampling**

At 50 dpf, all fish were euthanized using an ice slurry. All fish were patted dry and individually weighed. Three fish per treatment tank were randomly selected for gene expression analysis; the digestive tracts of these fish were dissected, flash frozen in liquid nitrogen, and stored at -80°C (Meyer et al. 2013). Five fish per treatment tank were randomly selected for histological analysis. Bodies were dissected at an angle below the anal fin and the anterior portion of the body was fixed in 10% neutral buffered formalin prior to histological processing (Sabaliauskas et al. 2006).

**Statistical Analysis**

Statistical analyses were performed using R software (R Foundation for Statistical Computing, Vienna, Austria, Version 4.2.1 “Funny-Looking Kid”). One-way ANOVA was used to assess differences in average individual weight, FCR, relative gene expression levels, and histological measurements among treatment groups. Normality was evaluated using the Shapiro Wilk's method. Levene’s tests were used to test for homogeneity of variances. All data mentioned above were normal with equal variances. Pairwise Tukey tests were used in the case of significance. Differences among treatment groups were considered significant at p values \< 0.05.

Percent survival was determined by dividing the final number of fish in each tank the initial number of fish in the tank and multiplying this value by 100. These values were analyzed using a Kruskal-Wallis one-way ANOVA due to non-normality. Feed conversion ratio (FCR) was also evaluated using a Kruskal-Wallis one-way ANOVA due to non-normality.

## Results

```{r, echo=FALSE, warning=FALSE, message=FALSE, results=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(car)
library(rstatix)
library(multcomp)

FRstudy <- read.csv('FRstudy.csv')
FRstudy_GE <-read.csv('FRstudy-GE.csv')

shapiro.test(FRstudy$Survival)
leveneTest(Survival~Treatment, data = FRstudy)

kruskal.test(Survival~Treatment, data = FRstudy)
```

Feeding regimen did not significantly affect survival (χ2 = 1.6, p = 0.6594).

```{r, echo=FALSE, fig.cap = 'Average individual weight (g) of fish at experimental termination (50 days-post-fertilization).'}
ggplot(data = FRstudy, aes(x = Treatment, y = w4_ind))+
  geom_boxplot()+
  theme_bw()+
  labs(x = 'Treatment', y = 'Final Average Weight (g)')
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results=FALSE}

shapiro.test(FRstudy$w4_ind)
leveneTest(w4_ind~Treatment, data=FRstudy)

weightaov <- aov(w4_ind~Treatment, data = FRstudy)
anova(weightaov)

```

There were no significant effects of dietary regimen on final individual weights (F3,8 = 2.1363, p = 0.1751). Numerically, the FM-SBM group had the highest weight per fish followed by FM/SBM, FM-FM, and SBM-SBM (Figure 1).

```{r, echo=FALSE, fig.cap = 'Average feed conversion ratio (FCR) after 26 days of feeding.'}
ggplot(data = FRstudy, aes(x = Treatment, y = FCR))+
  geom_boxplot()+
  theme_bw()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, results=FALSE}

shapiro.test(FRstudy$FCR)
leveneTest(FCR~Treatment, data=FRstudy)

FCRaov <- aov(FCR~Treatment, data = FRstudy)
anova(FCRaov)

kruskal.test(FCR~Treatment, data = FRstudy)

```

Feed conversion ratio (FCR) was not significantly different between treatment groups (X2 = 6.5897, df = 3, p = 0.08619; Figure 2).

```{r, echo=FALSE, warning=FALSE, message=FALSE, results=FALSE}
# il1b dct

shapiro.test(FRstudy_GE$il1b)
leveneTest(il1b~Treatment, data = FRstudy_GE)

il1b <-aov(il1b~Treatment, data = FRstudy_GE)
anova(il1b)

# tnfa dct

shapiro.test(FRstudy_GE$tnfa)
leveneTest(tnfa~Treatment, data = FRstudy_GE)

tnfa <-aov(tnfa~Treatment, data = FRstudy_GE)
anova(tnfa)

TukeyHSD(tnfa)

# mmp9 dct

shapiro.test(FRstudy_GE$mmp9)
leveneTest(mmp9~Treatment, data = FRstudy_GE)

mmp9 <-aov(mmp9~Treatment, data = FRstudy_GE)
anova(mmp9)

# il10

shapiro.test(FRstudy_GE$il10)
leveneTest(il10~Treatment, data = FRstudy_GE)

il10 <-aov(il10~Treatment, data = FRstudy_GE)
anova(il10)

```

```{r echo=FALSE, warning=FALSE, message=FALSE, results=FALSE}
inflammation <- FRstudy_GE %>% 
  dplyr::select(Treatment,il1b_FC:il10_FC) %>% 
  pivot_longer(cols = il1b_FC:il10_FC) %>% 
  group_by(Treatment, name) %>% 
  summarize(fold_change = mean(value), sd = sd(value), se = sd/sqrt(3))

labels <- c(
  il1b_FC = 'il1b',
  tnfa_FC = 'tnfa',
  mmp9_FC = 'mmp9',
  il10_FC = 'il10'
)
```

```{r graph, echo=FALSE, fig.cap = 'Expression of inflammation-related genes represented as average fold change in reference to the control group (FM-FM; n = 3). Error bars represent standard error of mean'}
ggplot(data = inflammation, mapping = aes(x = Treatment, y = fold_change, group = name))+
  geom_col(aes(fill = Treatment))+
  geom_errorbar(aes(ymin=fold_change-se, ymax=fold_change+se, width = 0.2))+
  theme_classic()+
  theme(axis.title.x=element_blank(),
      axis.text.x=element_blank(),
      axis.ticks.x = element_blank())+
  facet_wrap(vars(name), labeller = labeller(name = labels))+
  labs(x = 'Treatmeant', y = 'Fold Change')
```

Inflammation-related genes il-1b, tnfa, mmp9, and il-10 were examined. There was a significant effect of dietary treatment on the relative expression of tnfa (F3,8 = 6.9036, p = 0.0131). Pairwise Tukey tests revealed a significantly (p = 0.0090) higher expression of tnfa in the SBM-SBM group compared to the FM-FM group. There were no other significant effects of dietary treatment on the relative expression of inflammatory genes (p \>0.05; Figure 3).

```{r, echo=FALSE, warning=FALSE, message=FALSE, results=FALSE}

# ghrelin dct

shapiro.test(FRstudy_GE$ghrelin)
leveneTest(ghrelin~Treatment, data = FRstudy_GE)

ghrelin <-aov(ghrelin~Treatment, data = FRstudy_GE)
anova(ghrelin)

# leptin dct

shapiro.test(FRstudy_GE$leptin)
leveneTest(leptin~Treatment, data = FRstudy_GE)

leptin <-aov(leptin~Treatment, data = FRstudy_GE)
anova(leptin)

TukeyHSD(leptin)

# cck dct

shapiro.test(FRstudy_GE$cck)
leveneTest(cck~Treatment, data = FRstudy_GE)

cck <-aov(cck~Treatment, data = FRstudy_GE)
anova(cck)

```

```{r echo=FALSE, warning=FALSE, message=FALSE, results=FALSE}
appetite <- FRstudy_GE %>% 
  dplyr::select(Treatment,ghrelin_FC:cck_FC) %>% 
  pivot_longer(cols = ghrelin_FC:cck_FC) %>% 
  group_by(Treatment, name) %>% 
  summarize(fold_change = mean(value), sd = sd(value), se = sd/sqrt(3))

labels2 <- c(
  ghrelin_FC = 'ghrelin',
  leptin_FC = 'leptin',
  cck_FC = 'cck'
)
```

```{r, echo=FALSE, fig.cap = 'Expression of appetite-related genes represented as average fold change in reference to the control group (FM-FM; n = 3). Error bars represent standard error of mean.'}
ggplot(data = appetite, mapping = aes(x = Treatment, y = fold_change, group = name))+
  geom_col(aes(fill = Treatment))+
  geom_errorbar(aes(ymin=fold_change-se, ymax=fold_change+se, width = 0.2))+
  theme_classic()+
  theme(axis.title.x=element_blank(),
      axis.text.x=element_blank(),
      axis.ticks.x = element_blank())+
  facet_wrap(vars(name), labeller = labeller(name = labels2))+
  labs(x = 'Treatmeant', y = 'Fold Change')
```

Appetite-related genes ghrelin, leptin, and cck were also examined. There was a significant effect of treatment on the relative expression of leptin (p = 0.0150). Pairwise Tukey tests revealed significantly higher expression of leptin in the SBM-SBM group compared to the FM/SBM group (p = 0.0119). There were no significant effects on the relative expression of ghrelin or cck (p \> 0.05; Figure 4).

```{r echo=FALSE, warning=FALSE, message=FALSE, results=FALSE}
# histology

# proximal intestine
# villi length
shapiro.test(FRstudy$prox_length)
leveneTest(prox_length~Treatment, data = FRstudy)

prox_length <-aov(prox_length~Treatment, data = FRstudy)
anova(prox_length)

# villi width

shapiro.test(FRstudy$prox_width)
leveneTest(prox_width~Treatment, data = FRstudy)

prox_width <-aov(prox_width~Treatment, data = FRstudy)
anova(prox_width)

# villi length to width ratio

shapiro.test(FRstudy$prox_ratio)
leveneTest(prox_ratio~Treatment, data = FRstudy)

prox_ratio <-aov(prox_ratio~Treatment, data = FRstudy)
anova(prox_ratio)

# lamina propria

shapiro.test(FRstudy$prox_lp)
leveneTest(prox_lp~Treatment, data = FRstudy)

prox_lp <-aov(prox_lp~Treatment, data = FRstudy)
anova(prox_lp)

# muscularis

shapiro.test(FRstudy$prox_musc)
leveneTest(prox_musc~Treatment, data = FRstudy)

prox_musc <-aov(prox_musc~Treatment, data = FRstudy)
anova(prox_musc)

# middle intestine
# villi length
shapiro.test(FRstudy$mid_length)
leveneTest(mid_length~Treatment, data = FRstudy)

mid_length <-aov(mid_length~Treatment, data = FRstudy)
anova(mid_length)

# villi width

shapiro.test(FRstudy$mid_width)
leveneTest(mid_width~Treatment, data = FRstudy)

mid_width <-aov(mid_width~Treatment, data = FRstudy)
anova(mid_width)

# villi length to width ratio
shapiro.test(FRstudy$mid_ratio)
leveneTest(mid_ratio~Treatment, data = FRstudy)

mid_ratio <-aov(mid_ratio~Treatment, data = FRstudy)
anova(mid_ratio)

# lamina propria

shapiro.test(FRstudy$mid_lp)
leveneTest(mid_lp~Treatment, data = FRstudy)

mid_lp <-aov(mid_lp~Treatment, data = FRstudy)
anova(mid_lp)

# muscularis

shapiro.test(FRstudy$mid_musc)
leveneTest(mid_musc~Treatment, data = FRstudy)

mid_musc <-aov(mid_musc~Treatment, data = FRstudy)
anova(mid_musc)

# distal intestine
# villi length
shapiro.test(FRstudy$dis_length)
leveneTest(dis_length~Treatment, data = FRstudy)

dis_length <-aov(dis_length~Treatment, data = FRstudy)
anova(dis_length)

# villi width

shapiro.test(FRstudy$dis_width)
leveneTest(dis_width~Treatment, data = FRstudy)

dis_width <-aov(dis_width~Treatment, data = FRstudy)
anova(dis_width)

# villi length to width ratio
shapiro.test(FRstudy$dis_ratio)
leveneTest(dis_ratio~Treatment, data = FRstudy)

dis_ratio <-aov(dis_ratio~Treatment, data = FRstudy)
anova(dis_ratio)

# lamina propria

shapiro.test(FRstudy$dis_lp)
leveneTest(dis_lp~Treatment, data = FRstudy)

dis_lp <-aov(dis_lp~Treatment, data = FRstudy)
anova(dis_lp)

# muscularis

shapiro.test(FRstudy$dis_musc)
leveneTest(dis_musc~Treatment, data = FRstudy)

dis_musc <-aov(dis_musc~Treatment, data = FRstudy)
anova(dis_musc)

library(kableExtra)
```

```{r, echo = FALSE}

proximal <- FRstudy %>% 
  dplyr::select(Treatment,prox_length:prox_musc) %>% 
  group_by(Treatment) %>% 
  summarize(mean_length = mean(prox_length), sd_length = sd(prox_length), mean_width = mean(prox_width), sd_width = sd(prox_width), mean_ratio = mean(prox_ratio), sd_ratio = sd(prox_ratio), mean_lp = mean(prox_lp), sd_lp = sd(prox_lp), mean_musc = mean(prox_musc), sd_musc = sd(prox_musc))

middle <- FRstudy %>% 
  dplyr::select(Treatment,mid_length:mid_musc) %>% 
  group_by(Treatment) %>% 
  summarize(mean_length = mean(mid_length), sd_length = sd(mid_length), mean_width = mean(mid_width), sd_width = sd(mid_width), mean_ratio = mean(mid_ratio), sd_ratio = sd(mid_ratio), mean_lp = mean(mid_lp), sd_lp = sd(mid_lp), mean_musc = mean(mid_musc), sd_musc = sd(mid_musc))

distal <- FRstudy %>% 
  dplyr::select(Treatment,dis_length:dis_musc) %>% 
  group_by(Treatment) %>% 
  summarize(mean_length = mean(dis_length), sd_length = sd(dis_length), mean_width = mean(dis_width), sd_width = sd(dis_width), mean_ratio = mean(dis_ratio), sd_ratio = sd(dis_ratio), mean_lp = mean(dis_lp), sd_lp = sd(dis_lp), mean_musc = mean(dis_musc), sd_musc = sd(dis_musc))
```

Measurements of intestinal villi length, villi width, lamina propria, and muscularis ratio were taken from the proximal, middle, and distal portions of the intestine. There were no significant differences among treatment groups for any of the measurements (p > 0.05; Tables 1-3).

```{r, echo=FALSE}
kable(proximal, col.names = c('Treatment','Villi Length', 'sd', 'Villi Width', 'sd', 'Villi Length:Width', 'sd', 'Lamina Propria', 'sd', 'Muscularis', 'sd'), digits = 2, caption = 'Histological measurements (µm) of the proximal intestine. Values are presented as average and standard deviation. There were no significant differences between treatment groups (p > 0.05).') %>%
  add_header_above(c("Proximal" = 11)) %>% 
  kable_material() %>% 
  kable_styling(font_size = 8)

kable(middle, col.names = c('Treatment','Villi Length', 'sd', 'Villi Width', 'sd', 'Villi Length:Width', 'sd', 'Lamina Propria', 'sd', 'Muscularis', 'sd'), digits = 2, caption = 'Histological measurements (µm) of the middle intestine. Values are presented as average and standard deviation. There were no significant differences between treatment groups (p > 0.05).') %>%
  add_header_above(c("Middle" = 11)) %>% 
  kable_material() %>% 
  kable_styling(font_size = 8)

kable(distal, col.names = c('Treatment','Villi Length', 'sd', 'Villi Width', 'sd', 'Villi Length:Width', 'sd', 'Lamina Propria', 'sd', 'Muscularis', 'sd'), digits = 2, caption = 'Histological measurements (µm) of the distal intestine. Values are presented as average and standard deviation. There were no significant differences between treatment groups (p > 0.05).') %>%
  add_header_above(c("Distal" = 11)) %>% 
  kable_material() %>% 
  kable_styling(font_size = 8)
```

## Discussion

  Dietary plant protein sources, such as soybean meal, have been known to induce adverse effects, namely growth reduction and intestinal inflammation, in a broad spectrum of fish species (Fuentes-Appelgren et al. 2014, Ulloa et al. 2014). Due to the potential delay for the onset of SBMIE symptoms, this study sought to investigate alternating feeding regimens as a preventative method of mitigating diet-induced inflammation. 
  
**Growth Performance and Feed Utilization**

  Past studies have shown significant effects of SBM on weight gain and feed utilization in fish (Fontainhas-Fernandes et al. 1999, Zhou et al. 2005, de Francesco et al. 2004, Uran et al. 2009, Chen et al. 2011). These negative effects have been demonstrated in some studies of zebrafish (Micheloni et al. 2022, Hedrera et al. 2013), however, other studies have found that zebrafish are able to tolerate SBM with minimal consequences on their growth and health (Rehman et al. 2021, Dissinger et al. 2024). 
 
  While there were no significant differences in final average weight between treatment groups, numerically, the FM-SBM group had the highest weight per fish followed by FM/SBM, FM-FM, and SBM-SBM (Figure 1). Additionally, no significant differences in FCR was observed between groups, but the FM-FM group had the numerically lowest FCR followed by FM/SBM, FM-SBM, and finally the SBM-SBM group. Significantly increased relative expression of leptin, an anorexigenic hormone, in the SBM-SBM group in comparison to the FM/SBM group could point to some effect on appetite by SBM. Leptin has roles in both regulation of feed intake and control of body weight (Mania et al. 2017, Volkoff et al. 2005). Increased leptin levels are typically associated with decreased feed intake and increased energy expenditure (Volkoff et al. 2005). 
  

**Intestinal Health**

  There were no significant differences in histological measurements between treatment groups suggesting no morphological signs of inflammation were present in the intestine at the time of sampling. However, the SBM-SBM group had significantly higher relative expression of tnfa, a pro-inflammatory cytokine, compared to the FM-FM group. Additionally, the SBM-SBM group had numerically higher expression of all inflammation-related genes compared to the other groups. This could suggest onset of some inflammatory response potentially induced by the plant-based diet. 

  While inter-individual differences in immune response can be influenced by non-heritable factors such as environment and age, it is estimated that the genetic component can account for up to 40% of immune responses (Padovani et al. 2022). Strain-specific immunological profiles can exist; this was recently presented by Padovani et al. (2022) who found there to be significant differences in mortality and inflammatory responses of AB or TU larval zebrafish exposed to 2,4,6-Trinitrobenzenesulfonic acid (TNBS), a common agent used to induce inflammatory bowel disease (IBD). The use of pet-store derived, highly domesticated zebrafish in this study could provide some explanation for the lack of signficiant differences in weight, FCR, and expression of inflammation-related genes. 
 
  The results of this study could suggest a potential role of genetic background on the divergent results observed in zebrafish studies of SBM-induced enteritis (Valenzuela et al. 2021). Thus, future studies should investigate the use of different wild-type strain zebrafish to determine an appropriate nutritional model for investigating plant protein-based feeds and mitigation of intestinal inflammation. 

## References

Agboola JO, Chikwati EM, Hansen JØ, Kortner TM, Mydland LT, Krogdahl Å,  Djordjevic B, Schrama JW, Øverland M. A meta-analysis to determine factors associated with the severity of enteritis in Atlantic salmon (*Salmo salar*) fed soybean meal-based diets. Aquaculture. 2022a;555:738214. doi: 10.1016/j.aquaculture.2022.738214

Chen W, Ai Q, Mai K, et al. Effects of dietary soybean saponins on feed intake, growth performance, digestibility and intestinal structure in juvenile Japanese flounder (Paralichthys olivaceus). Aquaculture 2011;318(1–2):95–100

Dametto FS, Fior D, Idalencio R, Rosa JGS, Fagundes M, Marqueze A, Barreto RE,  Piato A, Barcellos LJG. Feeding regimen modulates zebrafish behavior. PeerJ. 2018:e5343. doi: 10.7717/peerj.5343

Dissinger A, Bowman J, Molinari GS, Kwasek K. Effects of oregano (*Origanum vulgare*) essential oil supplementation on growth performance of zebrafish (*Danio rerio*) fed a high-inclusion soybean meal diet. Zebrafish. 2024a;21(5):. doi: 10.1089/zeb.2024.0139

Dissinger A, Rimoldi S, Terova G, Kwasek K. Chronic social isolation affects feeding behavior of juvenile zebrafish (*Danio rerio*). PLoS ONE. 2024b;19(7):e0307967. doi: 10.1371/journal.pone.0307967

Fang L, Liang XF, Zhou Y, Guo XZ, He Y, Yi TL, Liu LW, Yuan XC, Tao YX. Programming effects of high-carbohydrate feeding of larvae on adult glucose metabolism in zebrafish, *Danio rerio*. British Journal of Nutrition. 2013;111(5):808-818. doi: 10.1017/S0007114513003243

Farias M, Certal AC. Different Feeds and Feeding Regimens have an Impact on Zebrafish Larval Rearing and Breeding Performance. International Journal of Marine Biology and Research. 1(1):1-18.

Fernandes H, Peres H, Carvalho AP. Dietary Protein Requirement During Juvenile  Growth of Zebrafish (*Danio rerio*). Zebrafish;13(6):548-555. doi: 10.1089/zeb.2016.1303

Fontainhas-Fernandes A, Gomes E, Reis-Henriques MA, Coimbra J. Replacement of Fish Meal by Plant Proteins in the Diet of Nile Tilapia: Digestibility and Growth Performance. Aquacult. Int. 1999;7:57–67.

de Francesco M, Parisi G, Médale F, Lupi P, Kaushik SJ, Poli BM. Effect of long-term feeding with a plant protein mixture based diet on growth and body/fillet quality traits of large rainbow trout (Oncorhynchus mykiss). Aquaculture. 2004;236(1–4):413–29.

Francis G, Makkar HPS, Becker K. Antinutritional factors present in plant-derived alternate fish feed ingredients and their effects in fish. Aquaculture. 2001;199:197-227. doi: 10.1016/S0044-8486(01)00526-9

Fuentes-Appelgren P, Opazo R, Barros L, Feijoo CG, Urzua V, Romero J. Effect of the Dietary Inclusion of Soybean Components on the Innate Immune System in Zebrafish. Zebrafish. 2014;11(1):41-49. doi: 10.1089/zeb.2013.0934

Fuentes-Quesada JP, Viana MT, Rombenso AN, Guerrero-Rentería Y, Nomura-Solís M, Gomez-Calle V, Lazo JP, Mata-Sotres JA. Enteritis induction by soybean meal in  Totoaba macdonaldi diets: Effects on growth performance, digestive capacity,    immune response and distal integrity. Aquaculture. 2018;495:78-89. doi: 10.1016/j.aquaculture.2018.05.025

Gonzales JM, Law SHW. Feed and Feeding Regime Affect Growth Rate and Gonadosomatic Index of Adult Zebrafish (*Danio rerio*). Zebrafish. 2013;10. doi: 10.1089/zeb.2013.0891

Gu M, Bai N, Zhang Y, Krogdahl Å. Soybean meal induces enteritis in turbot *Scophthalmus maximus* at high supplementation levels. Aquaculture. 2016;464:286-295. doi: 10.1016/j.aquaculture.2016.06.035

Hedrera MI, Galdames JA, Jimenez-Reyes MF, Reyes AE, Avedano-Herrera, R, Romero J, Feijoo CG. Soybean Meal Influences Intestinal Inflammation in Zebrafish Larvae. PLOS ONE. 2013;8(7):e69983.

Hua K, Cobcroft JM, Cole A, Condon K, Jerry DR, Mangott A, Praeger C, Vucko MJ, Zeng C, Zenger K, Strugnell JM. The Future of Aquatic Protein: Implications for Protein Sources in Aquaculture Diets. One Earth. 2019;1(3):316-329. doi: 10.1016/j.oneear.2019.10.018

Krogdahl Å, Penn M, Thorsen J, Refstie S, Bakke AM. Important antinutrients in  plant feedstuffs for aquaculture: an update on recent findings regarding  responses in salmonids. Aquaculture Research. 2010;41:333-344. doi: 10.1111/j.1365-2109.2009.02426.x

Kwasek K, Wojno M, Iannini F, McCracken VJ, Molinari GS, Terova G. Nutritional  programming improves dietary plant protein utilization in zebrafish *Danio rerio*. PLoS ONE. 2020;15(3):e0225917. doi: 10.1371/journal.pone.0225917

Kwasek K, Patula S, Wojno M, Oliaro F, Cabay C, Pinnell LJ. Does Exposure of    Broodstock to Dietary Soybean Meal Affect Its Utilization in the Offspring of   Zebrafish (*Danio rerio*)? Animals. 2022;12(2):1475. doi: 10.3390/ani12121475

Lawrence C, Best J, James A, Maloney K. The effects of feeding frequency on growth and reproduction in zebrafish (*Danio rerio*). Aquaculture. 2012;368-369:102-108. doi: 10.1016/j.aquaculture.2012.09.022 

Lin S, Luo L. Effects of different levels of soybean meal inclusion in replacement for fish meal on growth, digestive enzymes and transaminase activities in practical diets for juvenile tilapia, *Oreochromis niloticus* × *O. aureus*. Animal Feed and Science Technology 2011;168(1-2):80-87. doi: 10.1016/j.anifeedsci.2011.03.012

Makkar, H.P.S., 1993. Antinutritional factors in foods for livestock. In: Gill, M., Owen, E., Pollot, G.E.,Lawrence, T.L.J. Eds. , Animal Production in Developing Countries. Occasional publication No. 16. Ž .British Society of Animal Production, pp. 69–85.

Mania, M. et al. Expression and distribution of leptin and its receptors in the digestive tract of DIO (diet-induced obese) zebrafish. Anzeiger 212, 37-47. doi: 10.1016/j.aanat.2017.03.005

Meyer BM, Froehlich JM, Galt NJ, Biga PR. Inbred strains of zebrafish exhibit variation in growth performance and myostatin expression following fasting. Comparative Biochemistry and Physiology, A. 2013;164:1-9. doi: 10.1016/j.cbpa.2012.10.004

Micheloni G, Carnovali M, Millefanti G, Rizzetto M, Moretti V, Montalbano G, Acquati F, Giaroni C, Valli R, Costantino L, Ferrara F, Banfi G, Mariotti M, Porta G. Soy diet induces intestinal inflammation in adult Zebrafish: Role of OTX and P53 family. International Journal of Experimental Pathology. 2021;103:13-22. doi: 10.1111/iep.12420

National Research Council. Nutrient requirements of fish and shrimp. National   academies press, 2011.

Oliva-Teles A, Enes P, Peres H. Replacing fishmeal and fish oil in industrial   aquafeeds for carnivorous fish. Feed and Feeding Practices in Aquaculture. 2015:203-233. doi: 10.1016/B978-0-08-100506-4.00008-8

Padovani BN, do Amaral MA, Fénero CM, Paredes LC, de Barros GJB, Xavier IK, Hiyane MI, Ghirotto B, Feijóo CG, Câmara NOS, Takiishi T. Different wild type strains of zebrafish show divergent susceptibility to TNBS-induced intestinal inflammation displaying distinct immune cell profiles. Current Research in Immunology 2022;3:13-22. doi: 10.1016/j.crimmu.2021.12.003

Park H, Weier S, Razvi F, Peña PA, Sims NA, Lowell J, Hungate C, Kissinger K,   Key G, Fraser P, Napier JA, Cahoon EB, Clemente TE. Towards the development of  a sustainable soya bean-based feedstock for aquaculture. Plat Biotechnology Journal. 2017;15:227-236. doi: 10.1111/pbi.12608

Peng KS, Wu N, Cui ZW, Zhang XY, Lu XB, Wang ZX, Checn DD, Zhang YA. Effect of  the complete replacement of dietary fish meal by soybean meal on histopathology and immune response of the hindgut in grass carp  (*Ctenopharyngodon idellus*). Veterinary Immunology and Immunopathology. 2020;221:110009. doi: 10.1016/j.vetimm.2020.110009

Rehman S, Gora AH, Siriyappagouder P, Brugman S, Fernandes JM, Dias J, Kiron V. Zebrafish intestinal transcriptome highlights subdued inflammatory responses to dietary soya bean and efficacy of yeast β-glucan. Journal of Fish Diseases. 2021;44(10):1481-1668. doi: 10.1111/jfd.13484

Sabaliauskas NA, Foutz CA, Mest JR, Budgeon LR, Sidor AT, Gershenson JA, Joshi  SB, Cheng KC. High-throughput zebrafish histology. Methods. 2006;39(3):246-254. doi: 10.1016/j.ymeth.2006.03.001 

Sales J. The effect of fish meal replacement by soyabean products on fish growth: a meta-analysis. British Journal of Nutrition. 2009;102:1709-1722. doi: 10.1017/S0007114509991279

Ulloa PE, Medrano JF, Feijoo CG. Zebrafish as animal model for aquaculture nutrition research. Genetics. 2014;5(313):1-6. doi: 10.3389/fgene.2014.0031

Urán PA, Schrama JW, Rombout JHWM, Taverne-Thiele JJ, Obach A, Koppe W, Verreth JAJ. Time-related changes of the intestinal morphology of Atlantic salmon, *Salmo salar* L., at two different soybean meal inclusion levels. Journal of  Fish Diseases. 2009;32(9):733-744. doi: 10.1111/j.1365-2761.2009.01049.x

Valenzuela L, Pacheco S, Rincón G, Pavez L, Lam N, Hernández AJ, Dantagnan P, González F, Jilberto F, Ravanal MC, Ramos C, Garcia H, Araneda C, Ulloa PE. Intestinal Transcriptome Analysis Reveals Enrichment of Genes Associated with Immune and Lipid Mechanisms, Favoring Soybean Meal Tolerance in High-Growth Zebrafish (*Danio rerio*). Genes. 2021;12:700. doi: 10.3390/genes12050700

Volkoff, H. et al. Neuropeptides and the control of food intake in fish. Gen. Comp. Endocrinol. 142, 3-19 (2005). doi: 10.1016/j.ygcen.2004.11.001

Zhou QC, Mai KS, Tan BP, Liu YJ. Partial replacement of fishmeal by soybean meal in diets for juvenile cobia (*Rachycentron canadum*). Aquaculture Nutrition.2005;11(3):175-182. doi:10.1111/j.1365-2095.2005.00335.x
